Doppler Imaging Fun

### 2019-06-13 tex/gitlinks.tex
wooooot

\newcommand{\codelink}[1]{\href{https://github.com/user/repo/blob/a5fcbb15ba3e6d3c33496489dba346647c265f5d/tex/figures/#1.py}{\codeicon}\,\,}
\newcommand{\animlink}[1]{\href{https://github.com/user/repo/blob/a5fcbb15ba3e6d3c33496489dba346647c265f5d/tex/figures/#1.gif}{\animicon}\,\,}
\newcommand{\prooflink}[1]{\href{https://github.com/user/repo/blob/a5fcbb15ba3e6d3c33496489dba346647c265f5d/tex/proofs/#1.ipynb}{\raisebox{-0.1em}{\prooficon}}}

>>>>

\newcommand{\codelink}[1]{\href{https://github.com/user/repo/blob/d2fd278bdf8ebf22d2ce73093ca81d6bc0a4939b/tex/figures/#1.py}{\codeicon}\,\,}
\newcommand{\animlink}[1]{\href{https://github.com/user/repo/blob/d2fd278bdf8ebf22d2ce73093ca81d6bc0a4939b/tex/figures/#1.gif}{\animicon}\,\,}
\newcommand{\prooflink}[1]{\href{https://github.com/user/repo/blob/d2fd278bdf8ebf22d2ce73093ca81d6bc0a4939b/tex/proofs/#1.ipynb}{\raisebox{-0.1em}{\prooficon}}}

### 2019-06-13 tex/ms.tex
wooooot

\subsection{Taking the Fourier transform}
\label{sec:fourier}
%
Recall the definition of the Fourier transform of a function $f(\xi)$,
%
\begin{align}
    \label{eq:FT}
    \mathcal{F}\Big[f(\xi)\Big](k) 
    &\equiv
    \int_{-\infty}^\infty
        f(\xi)
        \mathrm{e}^{-2\pi i k \xi} \mathrm{d}\xi
    \nonumber \\
    &=
    \hat{f}(k)
    \quad ,
\end{align}
%
and the corresponding inverse Fourier transform,
%
\begin{align}
    \label{eq:IFT}
    \mathcal{F}^{-1}\Big[\hat{f}(k)\Big](\xi) 
    &\equiv
    \int_{-\infty}^\infty
        \hat{f}(k)
        \mathrm{e}^{2\pi i k \xi} \mathrm{d}k
    \nonumber \\
    &=
    f(\xi)
    \quad .
\end{align}
%
Consider what happens when we take the
Fourier transform of the spectrum $I(\xi, x, y)$ to obtain
a function $\hat{I}(k, x, y)$. A useful property of the
Fourier transform is that a translation of $I(\xi, x, y)$ by
an amount $\Delta\xi$ corresponds to a linear scaling
of $\hat{I}(k, x, y)$ by an amount $\mathrm{e}^{-2\pi i k \Delta\xi}$
\citep[e.g.,][]{Schoenstadt2006}.
Specifically, if the Doppler-shifted spectrum 
is given by
%
\begin{align}
    I(\xi, x, y) &= I_0\big(\xi + \alpha(x, y)\big) \quad,
\end{align}
%
its Fourier transform is equal to 
%
\begin{align}
    \label{eq:translation}
    \hat{I}(\xi, x, y) &= \mathrm{e}^{2\pi i k \alpha(x, y)}\hat{I_0}(k) \quad,
\end{align}
%
where $\hat{I_0}(k) \equiv \mathcal{F}\left[ I_0(\xi) \right]$.
%
Much as in \S\ref{sec:taylor}, we have de-coupled the spatial and spectral
terms, but this time without the need of a Taylor expansion. To obtain
the Fourier transform of the disk-integrated spectrum, $\hat{S}$, we may 
integrate Equation~(\ref{eq:translation})
over the visible disk $\mathcal{S}$ of the star:
%
\begin{align}
    \label{eq:integral_translation}
    \hat{S}(k) 
    &= 
        \hat{I_0}(k)\iint\limits_{\mathcal{S}(x, y)}
        \mathrm{e}^{2\pi i k \alpha(x, y)}
        \mathrm{d}\mathcal{S}(x, y) \nonumber \\
    &=
        \hat{I_0}(k)\iint\limits_{\mathcal{S}(x, y)}
        % MAGIC
        \left( \frac{1 - \beta(x, y)}{1 + \beta(x, y)} \right) ^ {\pi i k}
        % MAGIC
        \mathrm{d}\mathcal{S}(x, y) \nonumber \\
    &=
        \hat{I_0}(k) \hat{J}(k)
\end{align}
%
where we define
%
\begin{align}
    \hat{J}(k) \equiv 
        \iint\limits_{\mathcal{S}(x, y)}
        % MAGIC
        \left( \frac{1 - \beta(x, y)}{1 + \beta(x, y)} \right) ^ {\pi i k}
        % MAGIC
        \mathrm{d}\mathcal{S}(x, y)
\end{align}
%
to be the surface integral of the velocity field in Fourier space.
The disk-integrated spectrum is then obtained by applying the inverse Fourier
transform:
%
\begin{align}
    S(\xi) &= \mathcal{F}^{-1}\Big[ \hat{S}(k) \Big] \nonumber \\
           &= \mathcal{F}^{-1}\Big[ \hat{I_0}(k) \hat{J}(k) \Big] \\
           &= I_0(\xi) \circ J(\xi)
\end{align}
%
where the last step follows from the convolution theorem, 
with $J(\xi) \equiv \mathcal{F}^{-1}\left[ \hat{J}(k) \right]$ and $\circ$
denoting the convolution operator.

\subsubsection{Integration over the full stellar disk}
\label{sec:full_disk}
%
As an example, consider the case of an unocculted, uniform star of unit
radius with negligible limb darkening, rigidly rotating about an axis 
perpendicular to the line of sight with equatorial angular velocity 
$\omega_\mathrm{eq}$. The velocity profile everywhere on the surface is
%
\begin{align}
    \beta(x, y) &= \frac{\omega_\mathrm{eq}}{c}x \nonumber\\
                &\equiv \beta_\mathrm{eq} x 
                \quad .
\end{align}
%
The surface integral term may then be computed as
%
\begin{align}
    \hat{J}(k) 
    &=
        \int_{-1}^{1}
        \int_{-\sqrt{1 - x^2}}^{\sqrt{1 - x^2}}
        % MAGIC
        \left( 
            \frac{1 - \beta_\mathrm{eq} x}
                 {1 + \beta_\mathrm{eq} x} 
        \right) ^ {\pi i k}
        % MAGIC
        \mathrm{d}y
        \,
        \mathrm{d}x
        \nonumber \\[0.75em]
    &\approx
        \int_{-1}^{1}
        \int_{-\sqrt{1 - x^2}}^{\sqrt{1 - x^2}}
        % MAGIC
        \left( 
            1 - \beta_\mathrm{eq} x 
        \right) ^ {2\pi i k}
        % MAGIC
        \mathrm{d}y
        \,
        \mathrm{d}x
        \nonumber \\[0.75em]
    &=
        \pi \, _{2}F_{1}
            \left(
                -\pi i k, \, 
                \frac{1}{2} - \pi i k; \,
                2; \,
                \beta_\mathrm{eq}^2
            \right)
    \quad ,
\end{align}
%
where $_{2}F_{1}$ is the hypergeometric function. Note that
the approximation in the second line is valid when $\beta \ll 1$
(i.e., the non-relativistic case).

The equation for the observed disk-integrated spectrum is therefore
%
\begin{align}
    \label{eq:uniform}
    S(\xi) &= 
    \pi
    \mathcal{F}^{-1}
    \Big[
        \mathcal{F}(I_0)
        \,
        _{2}F_{1}
        \left(
            -\pi i k, \, 
            \frac{1}{2} - \pi i k; \,
            2; \,
            \beta_\mathrm{eq}^2
        \right)
    \Big]
\end{align}
%
Figure~\ref{fig:uniform} shows an example application of this equation
to the case of rotational broadening of a Gaussian absorption line.

@@

>>>>

\subsection{Taking the Fourier transform}
\label{sec:fourier}
%
Recall the definition of the Fourier transform of a function $f(\xi)$,
%
\begin{align}
    \label{eq:FT}
    \mathcal{F}\Big[f(\xi)\Big](k) 
    &\equiv
    \int_{-\infty}^\infty
        f(\xi)
        \mathrm{e}^{-2\pi i k \xi} \mathrm{d}\xi
    \nonumber \\
    &=
    \hat{f}(k)
    \quad ,
\end{align}
%
and the corresponding inverse Fourier transform,
%
\begin{align}
    \label{eq:IFT}
    \mathcal{F}^{-1}\Big[\hat{f}(k)\Big](\xi) 
    &\equiv
    \int_{-\infty}^\infty
        \hat{f}(k)
        \mathrm{e}^{2\pi i k \xi} \mathrm{d}k
    \nonumber \\
    &=
    f(\xi)
    \quad .
\end{align}
%
Consider what happens when we take the
Fourier transform of the spectrum $I(\xi, x, y)$ to obtain
a function $\hat{I}(k, x, y)$. A useful property of the
Fourier transform is that a translation of $I(\xi, x, y)$ by
an amount $\Delta\xi$ corresponds to a linear scaling
of $\hat{I}(k, x, y)$ by an amount $\mathrm{e}^{-2\pi i k \Delta\xi}$
\citep[e.g.,][]{Schoenstadt2006}.
Specifically, if the Doppler-shifted spectrum 
is given by
%
\begin{align}
    I(\xi, x, y) &= I_0\big(\xi + \alpha(x, y)\big) \quad,
\end{align}
%
its Fourier transform is equal to 
%
\begin{align}
    \label{eq:translation}
    \hat{I}(\xi, x, y) &= \mathrm{e}^{2\pi i k \alpha(x, y)}\hat{I_0}(k) \quad,
\end{align}
%
where $\hat{I_0}(k) \equiv \mathcal{F}\left[ I_0(\xi) \right]$.
%
Much as in \S\ref{sec:taylor}, we have de-coupled the spatial and spectral
terms, but this time without the need of a Taylor expansion. To obtain
the Fourier transform of the disk-integrated spectrum, $\hat{S}$, we may 
integrate Equation~(\ref{eq:translation})
over the visible disk $\mathcal{S}$ of the star:
%
\begin{align}
    \label{eq:integral_translation}
    \hat{S}(k) 
    &= 
        \hat{I_0}(k)\iint\limits_{\mathcal{S}(x, y)}
        \mathrm{e}^{2\pi i k \alpha(x, y)}
        \mathrm{d}\mathcal{S}(x, y) \nonumber \\
    &=
        \hat{I_0}(k)\iint\limits_{\mathcal{S}(x, y)}
        % MAGIC
        \left( \frac{1 - \beta(x, y)}{1 + \beta(x, y)} \right) ^ {\pi i k}
        % MAGIC
        \mathrm{d}\mathcal{S}(x, y) \nonumber \\
    &=
        \hat{I_0}(k) \hat{J}(k)
\end{align}
%
where we define
%
\begin{align}
    \label{eq:Jhat}
    \hat{J}(k) \equiv 
        \iint\limits_{\mathcal{S}(x, y)}
        % MAGIC
        \left( \frac{1 - \beta(x, y)}{1 + \beta(x, y)} \right) ^ {\pi i k}
        % MAGIC
        \mathrm{d}\mathcal{S}(x, y)
\end{align}
%
to be the surface integral of the velocity field in Fourier space.
The disk-integrated spectrum is then obtained by applying the inverse Fourier
transform:
%
\begin{align}
    S(\xi) &= \mathcal{F}^{-1}\Big[ \hat{S}(k) \Big] \nonumber \\
           \label{eq:S}
           &= \mathcal{F}^{-1}\Big[ \hat{I_0}(k) \hat{J}(k) \Big] \\
           \label{eq:S_conv}
           &= I_0(\xi) \circ J(\xi)
\end{align}
%
where the last step follows from the convolution theorem, 
with $J(\xi) \equiv \mathcal{F}^{-1}\left[ \hat{J}(k) \right]$ and $\circ$
denoting the convolution operator.

\subsubsection{Rotational line broadening}
\label{sec:line_broadening}
%
As an example, consider the case of an unocculted, uniform star of unit
radius with negligible limb darkening, rigidly rotating about an axis 
perpendicular to the line of sight with equatorial angular velocity 
$\omega_\mathrm{eq}$. The velocity profile everywhere on the surface is
%
\begin{align}
    \beta(x, y) &= \frac{\omega_\mathrm{eq}}{c}x \nonumber\\
                &\equiv \beta_\mathrm{eq} x 
                \quad .
\end{align}
%
The surface integral term may then be computed as
%
\begin{proof}{Fourier}
    \hat{J}(k) 
    &=
        \int_{-1}^{1}
        \int_{-\sqrt{1 - x^2}}^{\sqrt{1 - x^2}}
        % MAGIC
        \left( 
            \frac{1 - \beta_\mathrm{eq} x}
                 {1 + \beta_\mathrm{eq} x} 
        \right) ^ {\pi i k}
        % MAGIC
        \mathrm{d}y
        \,
        \mathrm{d}x
        \nonumber \\[0.75em]
    &\approx
        \int_{-1}^{1}
        \int_{-\sqrt{1 - x^2}}^{\sqrt{1 - x^2}}
        % MAGIC
        \left( 
            1 - \beta_\mathrm{eq} x 
        \right) ^ {2\pi i k}
        % MAGIC
        \mathrm{d}y
        \,
        \mathrm{d}x
        \nonumber \\[0.75em]
    &=
        \pi \, _{2}F_{1}
            \left(
                -\pi i k, \, 
                \frac{1}{2} - \pi i k; \,
                2; \,
                \beta_\mathrm{eq}^2
            \right)
    \quad ,
\end{proof}
%
where $_{2}F_{1}$ is the hypergeometric function. Note that
the approximation in the second line is valid when $\beta \ll 1$
(i.e., the non-relativistic case).

The equation for the observed disk-integrated spectrum is therefore
%
\begin{proof}{Fourier}
    \label{eq:uniform}
    S(\xi) &= 
    \pi
    \mathcal{F}^{-1}
    \Big[
        \mathcal{F}(I_0)
        \,
        _{2}F_{1}
        \left(
            -\pi i k, \, 
            \frac{1}{2} - \pi i k; \,
            2; \,
            \beta_\mathrm{eq}^2
        \right)
    \Big]
\end{proof}
%
Figure~\ref{fig:uniform} shows an example application of this equation
to the case of rotational broadening of a Gaussian absorption line.

@@

\subsubsection{Rossiter-McLaughlin Effect}
\label{sec:rmeffect}
%
Consider now the case where the integral is taken over only a portion
of the stellar disk, as in the case of an occultation by an orbiting planet
(or stellar companion). We can make use of Equation~(\ref{eq:S}) to
compute the time evolution of the observed spectrum during a transit
or occultation, i.e., the Rossiter-McLaughlin effect.
%
Working again in the non-relativistic limit, we may approximate
Equation~(\ref{eq:Jhat}) as
%
\begin{align}
    \label{eq:Jhat_approx}
    \hat{J}(k) \approx
        \iint\limits_{\mathcal{S}(x, y)}
        % MAGIC
        \left( 1 - \beta(x, y) \right) ^ {2\pi i k}
        % MAGIC
        \mathrm{d}\mathcal{S}(x, y)
\end{align}
%
We can further decompose this into its real and imaginary parts,
keeping only terms up to linear order in $\beta_\mathrm{eq}$:
%
\begin{align}
    \label{eq:Jhat_approx_re}
    \mathrm{Re}\left[\hat{J}(k)\right] 
    &\approx
        \iint\limits_{\mathcal{S}(x, y)}
        % MAGIC
        \cos \Big[ 2\pi k \ln\big(1 - \beta(x, y)\big) \Big]
        % MAGIC
        \mathrm{d}\mathcal{S}(x, y)
    \nonumber \\
    &\approx
        \iint\limits_{\mathcal{S}(x, y)}
        \mathrm{d}\mathcal{S}(x, y)
\end{align}
%
and
%
\begin{align}
    \label{eq:Jhat_approx_im}
    \mathrm{Im}\left[\hat{J}(k)\right] 
    &\approx
        \iint\limits_{\mathcal{S}(x, y)}
        % MAGIC
        \sin \Big( 2\pi k \ln\big(1 - \beta(x, y)\big) \Big)
        % MAGIC
        \mathrm{d}\mathcal{S}(x, y)
    \nonumber \\
    &\approx
        -2\pi k
        \iint\limits_{\mathcal{S}(x, y)}
        % MAGIC
        \beta(x, y)
        % MAGIC
        \mathrm{d}\mathcal{S}(x, y)
        \quad .
\end{align}
%
\cite{Luger2019} developed an analytic formalism to compute these 
integrals, provided $\beta(x, y)$ can be expressed in terms of
spherical harmonics. We will see in the following section that
$\beta(x, y)$ can be expressed exactly as the sum of
spherical harmonics of low degree in the case of a 
star with arbitrary inclination and obliquity and linear
differential rotation. For now, however, let us consider again
the simple case of a rigidly rotating star with rotation axis
perpendicular to the line of sight. In this case,
%
\begin{align}
    \mathrm{Re}\left[\hat{J}(k)\right] 
    &\approx \iint\limits_{\mathcal{S}(x, y)}\mathrm{d}\mathcal{S}(x, y)
    \nonumber \\
    &= \mathbf{s}_0(x_o, y_o, r_o)
\end{align}
%
and
%
\begin{align}
    \mathrm{Im}\left[\hat{J}(k)\right] 
    &\approx 
    -2\pi k \beta_\mathrm{eq}
    \iint\limits_{\mathcal{S}(x, y)}
    x
    \mathrm{d}\mathcal{S}(x, y)
    \nonumber \\
    &= -\pi k \beta_\mathrm{eq} \mathbf{s}_1(x_o, y_o, r_o)
\end{align}
%
